Measurement Bias from Address Aliasing

### 2014-08-15 paper.tex
Update haswell results for microkernel
Includes the updated LDM and L2 event cmask

% Large plot of cycle count for two 4K stack sections.
\pgfplotstableread{bin/micro-kernel-cycles.dat}{\stackoffsettable}
\begin{figure*}[t]
  \begin{tikzpicture}
    \begin{axis}[
        title = Haswell,
        width = \textwidth,
        height = 6cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Bias from environment size for microkernel. Measured average of 10 cycle count samples for 512 different environments. Spikes show aliasing case, occurring once for each 4K period.}
  \label{fig:envbias}
\end{figure*}

@@

% Detailed comparison between median and worst cases.
\begin{table}
  \tbl{Events with significant correlation to cycle count\label{tab:loopcorrelation}}{
    \pgfplotstabletypeset[
      int detect, % Output whole numbers for counter values
      col sep=comma,
      columns={Performance counter, Median, [index]3, [index]4},
      column type=r,
      columns/Performance counter/.style={
        string type, 
        column type=l,
        column type/.add={|}{},
        postproc cell content/.append code={
          \pgfkeysalso{@cell content=\perfctr{##1}}
        }
      },
      every head row/.style={
        output empty row,
        before row={\hline
          Performance counter & Median & Spike 1 & Spike 2 \\
        },
        after row=\hline\hline
      },
      every last row/.style={after row=\hline},
      every last column/.style={column type/.add={}{|}}
    ]{bin/micro-kernel-comparison.csv}
  }
  \begin{tabnote}
    \Note{Note:}{Some correlating performance events are omitted, for example bus-cycles which will naturally vary with total cycle count.}
  \end{tabnote}
\end{table}

We see the most extreme change from median to worst case in the number of alias events. 
If we plot the graph for address aliasing, we see that it is near zero everywhere and spikes at exactly the points we observe bias. 
The results show a high number of resource stalls and pending memory loads when the spikes occur, which is consistent with address aliasing issues.
On the other end we get a much lower number of reservation station (RS) stalls in the aliasing case, with a reduction from around 270,000 to 135,000.
The reservation station buffers micro-ops for scheduling to the execution units, and a stall event means that there are no free slots available~\cite[Table 19-2]{Volume3B}.
Fewer stalls in the aliasing case could indicate less contention on the reservation station.
This probably has to do with the overall \emph{decrease} in the number of micro-ops executed per port, as all but one of the in total 8 execution ports have fewer operations going in parallel.
Note that the number of micro-ops \emph{retired} overall does not change. %, but for some reason the micro-ops stay in the execution pipeline for a shorter time.
Bacause the memory disambiguation system does not handle address aliasing, the potential for out of order execution is limited~\cite[Page 2-20]{OptimizationManual}.
We assume that the tiny kernel size immediately prevents instructions from executing out of order, leading to the lower counts.
%Our assumption is that this is caused by the CPU being more conservative in the aliasing case, issuing fewer operations (speculatively) in parallel.
%With fewer instructions per cycle, there is also less pressure on the reservation station.

>>>>

% Large plot of cycle count for two 4K stack sections.
\pgfplotstableread{bin/microkernel-cycles-haswell.dat}{\stackoffsettable}
\begin{figure*}[t]
  \begin{tikzpicture}
    \begin{axis}[
        title = Haswell,
        width = \textwidth,
        height = 6cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Bias from environment size for microkernel. Measured average of 10 cycle count samples for 512 different environments. Spikes show aliasing case, occurring once for each 4K period.}
  \label{fig:envbias}
\end{figure*}

@@

% Detailed comparison between median and worst cases.
\begin{table}
  \tbl{Events with significant correlation to cycle count\label{tab:loopcorrelation}}{
    \pgfplotstabletypeset[
      int detect, % Output whole numbers for counter values
      col sep=comma,
      columns={Performance counter, Median, [index]3, [index]4},
      column type=r,
      columns/Performance counter/.style={
        string type, 
        column type=l,
        column type/.add={|}{},
        postproc cell content/.append code={
          \pgfkeysalso{@cell content=\perfctr{##1}}
        }
      },
      every head row/.style={
        output empty row,
        before row={\hline
          Performance counter & Median & Spike 1 & Spike 2 \\
        },
        after row=\hline\hline
      },
      every last row/.style={after row=\hline},
      every last column/.style={column type/.add={}{|}}
    ]{bin/microkernel-comparison-haswell.csv}
  }
  \begin{tabnote}
    \Note{Note:}{Some correlating performance events are omitted, for example bus-cycles which will naturally vary with total cycle count.}
  \end{tabnote}
\end{table}

We see the most extreme change from median to worst case in the number of alias events. 
If we plot the graph for address aliasing, we see that it is near zero everywhere and spikes at exactly the points we observe bias. 
The results show a high number of resource stalls and pending memory loads when the spikes occur, which is consistent with address aliasing issues.
On the other end we get a much lower number of reservation station (RS) stalls in the aliasing case, with a reduction from around 272,000 to 136,000.
The reservation station buffers micro-ops for scheduling to the execution units, and a stall event means that there are no free slots available~\cite[Table 19-2]{Volume3B}.
Fewer stalls in the aliasing case could indicate less contention on the reservation station.
This probably has to do with the overall \emph{decrease} in the number of micro-ops executed per port, as all but one of the in total 8 execution ports have fewer operations going in parallel.
Note that the number of micro-ops \emph{retired} overall does not change. %, but for some reason the micro-ops stay in the execution pipeline for a shorter time.
Bacause the memory disambiguation system does not handle address aliasing, the potential for out of order execution is limited~\cite[Page 2-20]{OptimizationManual}.
We assume that the tiny kernel size immediately prevents instructions from executing out of order, leading to the lower counts.
%Our assumption is that this is caused by the CPU being more conservative in the aliasing case, issuing fewer operations (speculatively) in parallel.
%With fewer instructions per cycle, there is also less pressure on the reservation station.
